### 2021-09-10 analysis/paper/supplementary_material.Rmd
Moved functions to script 3. Script 4 will be used to analyse the sites

---
title: "Supplementary material"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
institute:
  - UofO: University of One Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

```{r data_functions, eval = FALSE}
# Load in required functions and prepared data
library(here)
```

```{r data_functions, eval = TRUE, warning = FALSE}
# Load in required functions and prepared data
source(here("analysis/script/02script.R"))
source(here("analysis/script/03script.R"))
load(here("analysis/data/derived_data/01data.RData"))
load(here("analysis/data/derived_data/02data.RData"))
```

```{r , cache= = TRUE}

```

```{r demo-inline-code}

```

\newpage

# References

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies: